Improving Emotion Detection through Real-Time Translation of Text to ML Models Trained in Different Languages

### 2023-07-31 paper/appendix/dataset.tex
More Updates

\section{Appendix -- Dataset Details}
The following details in this appendix are for clarification purposes.
All examples in this appendix are created via PySpark using dataframes.

\subsection{Using PySpark to Load Data}
All example code was created using Jupyter Notebooks creating a PySpark session. The deails of setting up such session are below in the code example. All subsequent details are derived from the \texttt{sc}\footnote{sc = Spark Session}.
\begin{minted}{python}
# Setup Spark Session
# Spark Session Name: Qualyfing_Exam
from pyspark.sql import SparkSession

# Get Imports Needed
from pyspark.sql.functions import col, udf

# Get Datatypes needed for DataFrame manipulation
from pyspark.sql.types import IntegerType, StringType

# Setup Spark Session
sc = SparkSession \
        .builder \
        .master("local[*]") \
        .appName("Qualyfing_Exam") \
        .getOrCreate()
\end{minted}

\subsection{Details on English Dataset}
The English dataset based on tweets\footnote{Tweets from Twitter} are saved as a \texttt{csv}\footnote{csv = Comma Separated Values} file. The following are some simply details on the laod and processing of such file.
\begin{minted}{python}
# Read csv file to Saprk dataframe
df_english = sc.read.csv("data/english.csv", header=True, inferSchema=True)
# Group by a column and count the occurrences in df_english
count_english = df_english.groupBy('sentiment').count()

# Show the result for df_english
count_english.show()
+----------+-----+
| sentiment|count|
+----------+-----+
|   boredom|  179|
|      love| 3842|
|    relief| 1526|
|       fun| 1776|
|      hate| 1323|
|   neutral| 8638|
|     anger|  110|
| happiness| 5209|
|  surprise| 2187|
|   sadness| 5165|
|     worry| 8459|
|enthusiasm|  759|
|     empty|  827|
+----------+-----+

\end{minted}

\subsection{Details on German Dataset}
The German dataset was downloaded from the ETH's\footnote{ETH (German: Eidgenössische Technische Hochschule) or Swiss Federal Institute of Technology} servers via "link" as a \texttt{JSON} file.
\begin{minted}{python}
# Read JSON file to Saprk dataframe
df_german = sc.read.json("data/german.json")

# Group by a column and count the occurrences in df_german
count_german = df_german.groupBy('article_emotion').count()

# Show the result for df_german
count_german.show(truncate=0)
+----------------------------------+-----+
|article_emotion                   |count|
+----------------------------------+-----+
|[Ärger, Antizipation, Traurigkeit]|1    |
|[Überraschung, Vertrauen]         |14   |
|[Überraschung, Freude]            |15   |
|[Ärger, Vertrauen]                |7    |
|[Ekel, Traurigkeit]               |4    |
|[Keine, Antizipation]             |5    |
|[Vertrauen]                       |130  |
|[Freude]                          |58   |
|[Ärger, Antizipation]             |31   |
|[Ärger]                           |114  |
|[Überraschung]                    |154  |
|[Antizipation, Traurigkeit]       |15   |
|[Antizipation, Freude, Vertrauen] |2    |
|[Vertrauen, Angst]                |5    |
|[Traurigkeit]                     |101  |
|[Antizipation]                    |413  |
|[Traurigkeit, Vertrauen]          |7    |
|[Antizipation, Vertrauen]         |104  |
|[Ärger, Traurigkeit, Angst]       |1    |
|[Unklar]                          |314  |
+----------------------------------+-----+

\end{minted}

\clearpage

>>>>

\section{Appendix -- Dataset Details}
The following details in this appendix are for clarification purposes.
All examples in this appendix are created via Panda and PySpark using data frames.

\subsection{Translation Setup}
All example code was created using Jupyter Notebooks.
\begin{minted}[fontsize=\small]{python}
# *********************
# *** Setup Section ***
# *********************

import pandas as pd

# Emotion DataFrame
# This was predtermined by review of the emotion from English to German
emotion_key = {
    "boredom": "---",
    "love": "Vertrauen",
    "relief": "---",
    "fun": "---",
    "hate": "Ekel",
    "neutral": "Unklar",
    "anger": "Ärger",
    "happiness": "Freude",
    "surprise": "Überraschung",
    "sadness": "Traurigkeit",
    "worry": "Angst",
    "enthusiasm": "Antizipation",
    "empty ": "---",
    "---": "Keine"
}

# Create a DataFrame from the emotion_key dictionary
df_emotions = pd.DataFrame(emotion_key.items(), columns=['emotion_en', 'emotion_de'])

# *******************
# *** Translation ***
# *******************
from googletrans import Translator
import time

# This google API take a Sentence and converts to German
def translate(sentence, dest_lang):
    try:
        translator = Translator()
        translator.raise_Exception = True
        translation = translator.translate(sentence, dest=dest_lang)
        time.sleep(0.5)  # Add a 1-second delay
        return translation.text
    except Exception as e:
        print(f"Translation Error: {e}")
        return None
\end{minted}
\clearpage

\subsection{Details on English Dataset}
\label{appendix:dataset_english}
The English dataset is based on tweets\footnote{Tweets from Twitter} that are saved as a \texttt{csv}\footnote{csv = Comma Separated Values} file. The following code was use to filter and process the file for use in this project.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** English Data Preparations ***
# *********************************

import pandas as pd
from tqdm import tqdm

# Load English CSV File into a Dataframe
df_en = pd.read_csv("data/english.csv")
print(f'Count (raw): {len(df_en)}')

# Print Columns
print(f'English Source Columns: {df_en.columns}\n')

# Filter only labels that we matched with the German counter part JSON File
filter_values = ["love", "hate", "neutral", "anger", "happiness", "surprise", "sadness", "worry", "enthusiasm"]
df_en = df_en[df_en["sentiment"].isin(filter_values)]

# Find Empty Sentence Values
df_en = df_en[df_en["content"].notnull() & (df_en["content"].str.strip() != '')]

# Remove unnecessary column
df_en = df_en.drop("tweet_id", axis=1)

# Rename Columns
df_en = df_en.rename(columns = {'sentiment': 'emotion_en', 'content': 'sentence_en'})

# Merge German Emotions onto English
df_en = pd.merge(df_en, df_emotions, on='emotion_en', how='left')

# Add German Sentence Column
df_en["sentence_de"] = ""

# Randomly select 1500 rows
df_en = df_en.sample(n=1500, random_state=2023)

df_en.to_csv('./data/pd_en.csv', index=False)

# Iterate over the rows with tqdm to show the progress
for index, row in tqdm(df_en.iterrows(), total=df_en.shape[0]):
    # Call Translation
    sentence = translate(row["sentence_en"], 'de')
    # Save Sentence on Column
    df_en.at[index, 'sentence_de'] = sentence

df_en.to_csv('./data/pd_en_translated.csv', index=False)
\end{minted}
\clearpage

\subsection{Details on German Dataset}
\label{appendix:dataset_german}
The German dataset was downloaded from the ETH's\footnote{ETH (German: Eidgenössische Technische Hochschule) or Swiss Federal Institute of Technology} servers via "link" as a \texttt{JSON} file. Since the the file was \texttt{JSON} we will need to filter and explode the file based on teh emotions that are attached to it.
\begin{minted}[fontsize=\small]{python}
# *********************************
# *** German Data Preparations ***
# *********************************

import pandas as pd
from tqdm import tqdm

# Load German JSON File into a DataFrame
df_de = pd.read_json("data/german.json", lines=True)
print(f'Count (raw): {len(df_de)}')

# Print Columns
print(f'German Source Columns: {df_de.columns}\n')

# Explode the "article_emotion" column to distinct rows
df_de = df_de.explode('article_emotion')

# Filter only labels that we matched with the German counter part JSON File
filter_values = ["Vertrauen","Ekel", "Unklar", "Ärger", "Freude", "Überraschung", "Traurigkeit", "Angst", "Antizipation"]
df_de = df_de[df_de["article_emotion"].isin(filter_values)]

# Find Empty Sentence Values
df_de = df_de[df_de["title"].notnull() & (df_de["title"].str.strip() != '')]

# Remove unnecessary columns
df_de = df_de.drop("article_id", axis=1)
df_de = df_de.drop("article_stance", axis=1)
df_de = df_de.drop("paragraphs", axis=1)
df_de = df_de.drop("source", axis=1)
df_de = df_de.drop("snippet", axis=1)

# Rename Columns
df_de = df_de.rename(columns={'title': 'sentence_de'})
df_de = df_de.rename(columns={'article_emotion': 'emotion_de'})

# Merge Englsih Emotions onto German
df_de = pd.merge(df_de, df_emotions, on='emotion_de', how='left')

# Add English Sentence Column
df_de["sentence_en"] = ""

# Randomly select 1500 rows
df_de = df_de.sample(n=1500, random_state=2023)

df_de.to_csv('./data/pd_de.csv', index=False)

# Iterate over the rows with tqdm to show the progress
for index, row in tqdm(df_de.iterrows(), total=df_de.shape[0]):
    # Call Translation
    sentence = translate(row["sentence_de"], 'en')
    # Save Sentence on Column
    df_de.at[index, 'sentence_en'] = sentence

df_de.to_csv('./data/pd_de_translated.csv', index=False)
\end{minted}
\clearpage

### 2023-07-31 paper/main.tex
More Updates

@@

\begin{abstract}
This Qualifying Exam\footnote{PhD students of MTSU Computational and Data  Science program are required to complete a Qualifying Exam in their first year in the program.} research paper investigates the potential of improving Emotion Detection (ED) through real-time translation of text to multiple Machine Learning (ML) models trained in different languages. The research tries to address the challenges posed by the lack of comprehensive labeled datasets and language fragmentation in ED analysis.

By extending an original language dataset with a translated dataset from another language and utilizing parallel processing with PySpark, this research project aims to achieve higher prediction rates in ED. The results presented both orally to the MTSU's Computational Data \& Science Committee and within this paper demonstrate the effectiveness of such an approach and how the real-time translation technique can enhance emotion detection capabilities.
\end{abstract}
\clearpage

@@

Emotion detection has gained significant interest in recent years as a means to understand human behavior and communication better. By analyzing text or speech, these detection processes can accurately identify and classify emotions expressed by individuals or groups. By employing a variety of ED processes, including natural language processing and machine learning, we can detect emotions with high precision. These ED systems find applications in diverse fields, such as sentiment analysis, customer feedback analysis, and mental health monitoring, contributing to a deeper understanding of human emotions and their impact on various aspects of life.

\subsection{Background on Human Emotion Theoretical Framework}

The concept of the eight main emotions can vary depending on the theoretical framework or model being considered. One well-known model is the Plutchik's Wheel of Emotions Figure \ref{fig:plutchik-wheel}, which proposes eight primary emotions and their contrasting pairs. Below are the eight primary emotions according to the Plutchik's Wheel of Emotions \cite{Tromp}, a detailed reference of emotions used in this research can be found in Table \ref{table:dataset_labels} on page~\pageref{table:dataset_labels}.

\begin{itemize}
\item Joy (Fun): A feeling of happiness, contentment, or delight.
\item Sadness: A state of unhappiness, sorrow, or grief.
\item Anger: An intense emotional response associated with feelings of displeasure, frustration, or hostility.
\item Worry: An emotional response triggered by perceived threats or danger, often associated with anxiety or terror.
\item Surprise: A sudden and unexpected reaction to something unusual or startling.
\item Disgust (Hate): A strong aversion or revulsion towards something unpleasant, offensive, or repulsive.
\item Trust: A sense of reliance, confidence, or faith in someone or something.
\item Anticipation: A feeling of excited expectation or eagerness towards future events.
\end{itemize}

@@

It's important to note that emotions are complex and multifaceted, and different models may propose alternative categorizations or variations in the number of primary emotions. However for the scope of this research project the most common 8 emotions are used as described above. Both the English \cite{english-dataset-twitter} and German \cite{german-dataset-cheese} datasets being used in this research matched to the 8 listed above. Details on the linkage of the two datasets is described in Section \ref{sec:data-procurement-process} with an emphasis on the Table \ref{table:dataset_labels}'s detailed linkage.

\subsection{Background and significance of Emotion Detection (ED) analysis in Machine Learning models}
Emotion Detection (ED) analysis in Machine Learning (ML) models has gained significant attention due to its potential applications in various fields, including customer sentiment analysis, stance detection in regard to a specific targets such as news and voting\cite{mascarell-etal-2021-stance}, mental health monitoring\cite{Colonnello}, and human-computer interactions to just name a few.

By accurately identifying and understanding emotions from text data, ML models can assist in improving user experiences, decision-making processes, and overall human-machine interactions in a positive manner\cite{Colonnello, mascarell-etal-2021-stance}. As an example, the primary goals of chatbots are entertainment, social contact, and novelty interaction, with a strong emphasis on productivity. In chatbots, there has been a tremendous requirement to simulate human-like characteristics and behavior.With the goal in mind of improving ML models this research paper came to be \cite{emotion-detection-literature-review}. These chatbots also improve customer engagement by offering friendliness, comfortness, flexibility, and efficient assistance. Chatbots should provide customers with more engaging responses, directly addressing their issues. Users perceive chatbots as companions rather than simple assistants. The majority of user requests are emotional rather than informative. Chatbots gained the capacity to respond emotionally to customers due to machine learning and sentiment analysis evolution\cite{chatbot-cognitive-awareness}.

@@

Due to the before mentioned wide scope of uses for ED this field of research and the work to advance it provides significance and importance in enabling machines to comprehend and respond appropriately to human emotions. This in turn contributes to the advancement of artificial intelligence, natural language processing and in many cases the utilization of text-to-speech applications most all of us in today's society use and rely on.

\subsection{ED challenges, including lack of comprehensive labeled datasets and language fragmentation}
ED is growing field, however unlike Sentiment Analysis (SA) the availability of large datasets for training purposes of ML models is much smaller \cite{ACLU-ED-Data}. The primary reason of the lack of data stems from the emotion detection requiring primarily supervised learning dataset, which require time and effort to procure, and mostly created by humans. To compound on the issue of the lack of data; the many datasets that are available are in multiple languages.

@@

\subsection{Objectives \& Scope of the Research}
The project is comprised of three main points / goals which are listed in following sub-sections with details in each point as follows:

\subsubsection{Data Collection / Translation / ML Model Training}

Once the two datasets have been translated, we now extend the original language dataset with the translated one.By this approach we are in essence extending the original datasets to create a larger one wiht the foreign language data. With both extended datasets we can now train our ML models for English and German.

\subsubsection{PySpark Application Development for Translation \& Prediction}
A program will be built in Jupiter Notebooks that uses translations APIs to convert an English sentence to German and the same if a German sentence is fed to English. Both sentences are then processed in real-time on both the English and German models for the Emotion Detection and will be compared to each other's output. The average of the prediction will be used for final prediction of the supplied sentence and the emotion that was detected.

\section{Literature Review}

\subsection{Emotion Detection Importance, Data Limitations, Diverse Populations}
Emotion Detection (ED) holds significant importance in various domains, including psychology, social sciences, and human-computer interaction. It enables the analysis of human emotions at scale, providing valuable insights into individual and collective emotional states. However, the field faces limitations due to the scarcity of comprehensive labeled datasets. The lack of diverse and large-scale data hinders the development and training of accurate ED models, limiting their effectiveness and general ability across different contexts and populations.

In recent research, efforts have been made to address these limitations by exploring new techniques for emotion detection. For instance, a study by Shalini Kapoor, Tarun Kumar \cite{KAPOOR2023120882} proposes analyzing the entire spectrum of the language source to predict emotional changes.

Overcoming the scarcity of labeled datasets and embracing innovative approaches like the before mentioned study by Shalini Kapoor, Tarun Kumar \cite{KAPOOR2023120882} and their analysis can contribute to the advancement of ED research. Such advancements hold the potential to enhance the accuracy and applicability of emotion detection models, enabling a deeper understanding of human emotions and more effective responses in various domains.

Emotions can differ across age groups, genders, cultures, and languages. Including data from diverse populations helps in developing inclusive and culturally sensitive ED models. It ensures that the models are not biased towards specific groups and can accurately detect emotions in a wide range of individuals.

\subsection{Challenges by the lack of Labeled Datasets and the Fragmentation caused by Different Languages}
Unlike images language operate very differently, therefore the lack of labeled datasets poses a significant challenge in text-based Emotion Detection (ED) analysis as it limits the availability of reliable training data for ML models. This scarcity hampers the model's ability to learn and generalize emotions effectively. Additionally, the fragmentation caused by different languages further exacerbates the issue, as it reduces the size and diversity of data available for training, resulting in limited cross-lingual generalization and potentially biased models. Overcoming these challenges is crucial for advancing ED research and improving the accuracy and applicability of emotion detection models across various languages and cultures.

@@

\subsection{Data procurement process, including obtaining ED datasets in English and German}
\label{sec:data-procurement-process}
The data procurement was relatively straight forward and in essence entailed using Google search terms for the Emotion Detection in English and German. Some of the main German data was obtained from the dataset built by ETH's Emotion and Stance Detection for German Text \cite{mascarell-etal-2021-stance} which also hosts a good website with their finds \footnote{https://mtc.ethz.ch/research/natural-language-processing/emotion-stance.html}. The English dataset was downloaded from Kaggle \footnote{https://www.kaggle.com/datasets/pashupatigupta/emotion-detection-from-text} based on Tweets collected in 2021. The English dataset contains over 38,000 tweets and their corresponding emotion labeled. Similarly the German contains over 12,000 sentences and their respective emotions associated.

@@

@@

\subsection{Building and training multiple ML models using the extended datasets and parallel processing approach with PySpark}
\ldots

>>>>

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{1.2in}}

@@

\begin{abstract}
This Qualifying Exam\footnote{PhD students of MTSU Computational and Data  Science program are required to complete a Qualifying Exam in their first year in the program.} research paper investigates the potential of improving Emotion Detection (ED) through real-time translation of text to multiple Machine Learning (ML) models trained in different languages (English and German). The research tries to address the challenges posed by the lack of comprehensive labeled datasets and language fragmentation in ED analysis.

By extending an original language dataset with a translated dataset, in the case of this paper English and German. And utilizing parallel processing with PySpark, this research project aims to achieve higher prediction rates in ED. The results presented both orally to the MTSU's Computational Data \& Science Committee and within this paper demonstrate the effectiveness of such an approach and how the real-time translation technique can enhance emotion detection capabilities.
\end{abstract}
\clearpage

@@

Emotion detection has gained significant interest in recent years as a means to understand human behavior and improve communication effectiveness. By analyzing text or speech, these detection processes can accurately identify and classify emotions expressed by individuals or groups, often times in real-time.

By employing a variety of ED processes, including natural language processing and machine learning, applications can detect emotions with high precision. These ED systems find applications in diverse fields, such as sentiment analysis, customer feedback analysis, voting and news stances, and mental health monitoring, which all aim at contributing to a deeper understanding of human emotions and their impact on various aspects of life.

\subsection{Human Emotion as a Theoretical Framework}

The concept and number of emotions can vary depending on the theoretical framework or model being considered. One well-known model is the Plutchik's Wheel of Emotions Figure \ref{fig:plutchik-wheel}, which proposes eight (8) primary emotions and their contrasting pairs. In this paper we will be focusing on the basic eight emotions and not on their sub-emotions.

Below are the eight primary emotions according to the Plutchik's Wheel of Emotions \cite{Tromp}, a detailed reference of emotions used in this research can be found in Table \ref{table:dataset_labels} on page~\pageref{table:dataset_labels}.

\begin{itemize}
\item Happiness: A feeling of Joy (Fun), contentment, or delight.
\item Sadness: A state of unhappiness, sorrow, or grief.
\item Anger: An intense emotional response associated with feelings of displeasure, frustration, or hostility.
\item Worry: An emotional response triggered by perceived threats or danger, often associated with anxiety or terror.
\item Surprise: A sudden and unexpected reaction to something unusual or startling.
\item Disgust and/or Hate: A strong aversion or revulsion towards something unpleasant, offensive, or repulsive.
\item Trust and/or Love: A sense of reliance, confidence, or faith in someone or something.
\item Enthusiasm: A feeling of excited expectation or eagerness towards future events.
\end{itemize}

@@

It's important to note that emotions are complex and multifaceted, and different models may propose alternative categorizations or variations in the number of primary emotions. However for the scope of this research project the most common eight (8) emotions are used as described above.

By aligning the emotional labels across the English \cite{english-dataset-twitter} and German \cite{german-dataset-cheese} datasets being used in this research all eight listed above were used. Details on the linkage of the two datasets is described in Section \ref{sec:data-procurement-process} with an emphasis on the Table \ref{table:dataset_labels}'s cross reference on page~\pageref{table:dataset_labels}.

\subsection{Significance of Emotion Detection (ED) in Machine Learning}
Emotion Detection (ED) analysis in Machine Learning (ML) models has gained significant attention due to its potential applications in various fields, including customer sentiment analysis, stance detection in regard to a specific targets such as news and voting\cite{mascarell-etal-2021-stance}, mental health monitoring\cite{Colonnello}, and human-computer interactions (chatbots\cite{chatbot-cognitive-awareness}) to just name a few.

By accurately identifying and understanding emotions from text data, ML applications can assist in improving user experiences, decision-making processes, and overall human-machine interactions in a positive manner\cite{Colonnello, mascarell-etal-2021-stance}. As an example, the primary goals of chatbots are entertainment, social contact, and novelty interaction, with a strong emphasis on productivity, scale-ability and cost reductions. In chatbots, there has been a immense requirement to simulate human-like characteristics and behavior during human-machine interactions\cite{emotion-detection-literature-review}. Per Kusal et al. \cite{chatbot-cognitive-awareness} these chatbots also improve customer engagement by offering friendliness, comfortness, flexibility, and efficient assistance. Chatbots should provide customers with more engaging responses, directly addressing their issues. The goal of deploying ED applications is for users perceive chatbots as companions rather than simple assistants. Per Adikari et al. the majority of user requests are emotional rather than informative. Chatbots gained the capacity to respond emotionally to customers due to machine learning and sentiment analysis evolution\cite{chatbot-cognitive-awareness}.

@@

Due to the before mentioned wide scope of uses for ED this field of research and the work to advance it provides significance and importance in enabling machines to comprehend and respond appropriately to human interactions. This in turn contributes to the advancement of artificial intelligence, natural language processing and in many cases the utilization of text-to-speech applications most all of us in today's society use and rely on \cite{ai-framework-detection-emotions}.

\subsection{ED challenges: Lack of Comprehensive Labeled Datasets and Language Fragmentation}
ED is an immensely growing field, however unlike Sentiment Analysis (SA) the availability of large datasets for training purposes of ML models is much smaller \cite{ACLU-ED-Data, ai-framework-detection-emotions}. The primary reason of the lack of data stems from the emotion detection data requiring primarily supervised learning dataset, which require time and effort to procure, and mostly created and labeled by humans. To compound on the issue of the lack of data; the many datasets that are available are in multiple languages. A significant amount of time took place to find the German labeled

@@

\subsection{Objectives \& Scope of the Research}

The research project aims at answering the following research questions:

\begin{itemize}
\item Can by translating English data to German and extending an original German dataset increase the predictability of the ED model?
\item Can by translating German data to English and extending an original English dataset increase the predictability of the ED model?
\item Can by translating in real-time an input to multiple languages improve the predictability based on the combined output of two model's results.
\end{itemize}

\section{Literature Review}
To ensure comprehensive coverage of the latest developments and research in ED for this research paper, consideration and review was only given to incorporating those research papers and online news articles that were published after the year 2018, thereby providing an up-to-date perspective on the subject matter.

\subsection{Emotion Detection Importance and Diverse Populations}
Emotion Detection (ED) holds significant importance in various domains, including psychology, social sciences, and human-computer interaction. It enables the analysis of human emotions at scale, providing valuable insights into individual and collective emotional states.

Furthermore, emotions can differ across age groups, genders, cultures, and languages. Including data from diverse populations helps in developing inclusive and culturally sensitive ED models. It ensures that the models are not biased towards specific groups and can accurately detect emotions in a wide range of individuals.

\subsection{Lack of Labeled Datasets and the Fragmentation caused by Different Languages}

The lack of diverse and large-scale data hinders the development and training of accurate ED models, limiting their effectiveness and general ability across different contexts and populations.

In recent research, efforts have been made to address these limitations by exploring new techniques for emotion detection. For instance, a study by Kapoor\cite{KAPOOR2023120882} et al. proposes analyzing the entire spectrum of the language source to predict emotional changes. This provided some help, but quickly established itself as time consuming and need of expertise in the subject matters. Overcoming the scarcity of labeled datasets and embracing innovative approaches like the before mentioned study by Kapoor\cite{KAPOOR2023120882} et al. and their analysis can contribute to the advancement of ED research. Such advancements hold the potential to enhance the accuracy and applicability of emotion detection models, enabling a deeper understanding of human emotions and more effective responses in various domains.

The paper by Mascarell\cite{mascarell-etal-2021-stance} et al. states that unlike images language operate very differently, therefore the lack of labeled datasets poses a significant challenge in text-based Emotion Detection (ED) analysis as it limits the availability of reliable training data for ML models. This scarcity hampers the model's ability to learn and generalize emotions effectively.

Additionally, Kusal et al. further stipulates \cite{kusal} that the fragmentation caused by different languages further exacerbates the issue, as it reduces the size and diversity of data available for training, resulting in limited cross-lingual generalization and potentially biased models. Overcoming these challenges is crucial for advancing ED research and improving the accuracy and applicability of emotion detection models across various languages and cultures.

@@

\subsection{Data procurement process of ED datasets in English and German}
\label{sec:data-procurement-process}
The data procurement was relatively straight forward and in essence entailed using Google search terms for the Emotion Detection in English and German. Some of the main German data was obtained from the dataset built by ETH's Emotion and Stance Detection for German Text \cite{mascarell-etal-2021-stance} which also hosts a good website with their findings \footnote{https://mtc.ethz.ch/research/natural-language-processing/emotion-stance.html}. The English dataset was downloaded from Kaggle \footnote{https://www.kaggle.com/datasets/pashupatigupta/emotion-detection-from-text} based on Tweets collected in 2021. The English dataset contains over 38,000 tweets and their corresponding emotion labeled. Similarly the German contains over 12,000 sentences and their respective emotions associated.

@@

Once the two datasets have been translated, we now extend the original language dataset with the translated one.By this approach we are in essence extending the original datasets to create a larger one with the foreign language data. With both extended datasets we can now train our ML models for English and German.

@@

Based on the above example the CSV and JSON datasets were individually loaded, parsed, and translated to their respective counter-part languages. Details of the application can be found in the Appendix \ref{appendix:dataset_english} for English and Appendix \ref{appendix:dataset_german} for the German. We decided to only use 1,500 row for each language so since the German (DE\footnote{DE = German}) dataset was not much larger than that. This meant that the English (EN\footnote{EN = Enlgish}) datasets was randomly paired down to 1,500 to match that of the German.

Each translated file now contains four (4) columns that are labeled as in Tables \ref{table:pd_de_translated} and \ref{table:pd_en_translated} with the exact same column naming.

\begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{File: \texttt{pd\textunderscore de\textunderscore translated.csv}} \\
    \hline
    sentence\textunderscore de &
    emotion\textunderscore de & 
    sentence\textunderscore en & 
    emotion\textunderscore en \\
    \hline
    Original Sentence (feature) &
    Original Emotion (label) &
    Sentence Translated to English & 
    Emotion set by Cross-Reference from German \\

    \hline
\end{tabular}
\caption{German CSV file structure after translation to English}
\label{table:pd_de_translated}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{File: \texttt{pd\textunderscore en\textunderscore translated.csv}} \\
    \hline
    sentence\textunderscore en &
    emotion\textunderscore en & 
    sentence\textunderscore de & 
    emotion\textunderscore de \\
    \hline
    Original Sentence (feature) &
    Original Emotion (label) &
    Sentence Translated to German & 
    Emotion set by Cross-Reference from English \\

    \hline
\end{tabular}
\caption{English CSV File structure after translation to German}
\label{table:pd_en_translated}
\end{table}

\subsection{Training Multiple ML models using the Extended Datasets with PySpark}

 Both sentences are then processed in real-time on both the English and German models for the Emotion Detection and will be compared to each other's output. The average of the prediction will be used for final prediction of the supplied sentence and the emotion that was detected.

 \begin{table}[h!]
\centering
\begin{tabular}{ | L | L | L | L | }
    \hline
    \multicolumn{4}{|c|}{ML Models } \\
    \hline
    Model & 
    Data &
    Rows for Training & 
    Rows for Testing  \\
    \hline
    A &
    English Original &
    1200 & 
    300 (20\%) \\
    \hline
    B &
    English Extended By German &
    2700 & 
    300 same as Model "A" \\
    \hline
    C &
    German Original &
    1200 & 
    300 (20\%) \\
    \hline
    D &
    German Extended By English &
    2700 & 
    300 same as Model "C" \\

    \hline
\end{tabular}
\caption{ML Models and their respective settings}
\label{table:ml_models}
\end{table}
